Legacy Data Confounds Modern Genomics Studies

### 2019-07-12 PlottingScripts/ProcessPlotData.Rmd

metaData
#save to file
saveRDS(metaData,
            paste0(DataPath,
                   'metaData.rds'))
```
##Overlap of Significant SNPs from Single Population test
```{r}
phase1 <- c('YRI','LWK','CHB','CHD','JPT','CEU','LWK')
#read single population regressions
Reg <- fread(paste0(DataPath,
                    'AllRegressionsOver10.txt'), 
             col.names = c('Chr','Pos',
                           'Population','deviance'), 
             colClasses = c('numeric', 'numeric', 
                            'character', 'numeric'))

strictMask <- fread(paste0(DataPath,
                           '1kGP_StrictMask.bed'),
            col.names = c('Chr','Pos','Pos.1','rsID'))
#remove positions that don't pass strict mask
Reg <- anti_join(Reg,
                 strictMask,
                 by = c('Chr','Pos'))

#Get P-Values and significant sites
Reg$plog10 <- -log10(pchisq(Reg$deviance, 1, 
                            lower.tail = FALSE))
Reg<-Reg[which(Reg$plog10 >= 6),]
#Count number of populations where a variant is significant
NumPop <- as.data.frame(table(Reg$Chr,Reg$Pos))
NumPop <- NumPop[which(NumPop$Freq>1),]
names(NumPop) <- c('Chr','Pos', 'SigHits')
NumPop$Chr<-as.numeric(as.character(NumPop$Chr)) 
NumPop$Pos<-as.numeric(as.character(NumPop$Pos))

m <- merge(Reg,
           NumPop, 
           by = c('Chr','Pos'))
#count number of significant sites per population
n <- as.data.frame(table(m$Pop))
#put it all together for plotting
overLap <- unique(merge(m,n,
                        by.x = 'Population',
                        by.y = 'Var1'))
#include number of significant sites in titles
overLap$title = paste0(overLap$Pop,
                       ' : ', 
                       overLap$Freq)
#add asterix for Phase 1 pops
Phase1Title <- paste('*',
                     overLap[which(overLap$Pop 
                                   %in% phase1),]$title)
overLap[which(overLap$Pop
              %in% phase1),]$title <- Phase1Title
#set order by SigHits, Chromosome and Position
overLap <- overLap[with(overLap, 
                        order(-SigHits,
                              Chr,
                              Pos)),]
overLap$Pos <- factor(overLap$Pos, 
                      levels = unique(overLap$Pos))
#this is for keeping track of the number of variants
overLap$nrow <- seq(length = nrow(overLap))
#this is for discrete point size
overLap$Pcat <- 4
overLap[which(overLap$plog10 < 8),]$Pcat <- 2
overLap[which((overLap$plog10 > 8) & 
              (overLap$plog10 < 10)),]$Pcat <- 3

overLap

#save to file
saveRDS(overLap,
            paste0(DataPath,
                   'overLap.rds'))
```

>>>>

metaData
#save to file
saveRDS(metaData,
            paste0(DataPath,
                   'metaData.rds'))
```
##Overlap of Significant SNPs from Single Population test
```{r}
strictMask <- fread(paste0(DataPath,
                           '1kGP_StrictMask.bed'),
            col.names = c('Chr','Pos','Pos.1','rsID'))

phase1 <- c('YRI','LWK','CHB','CHD','JPT','CEU','LWK')
#read single population regressions
Reg <- fread(paste0(DataPath,
                    'AllRegressionsOver10.txt'), 
             col.names = c('Chr','Pos',
                           'Population','deviance'), 
             colClasses = c('numeric', 'numeric', 
                            'character', 'numeric'))

# positions that don't pass strict mask
Reg_not_strict <- inner_join(Reg,
                 strictMask,
                 by = c('Chr','Pos')) %>%
  select('Chr','Pos','Population','deviance') %>%
  mutate(strict = "fail")

#remove positions that don't pass strict mask
Reg_strict <- anti_join(Reg,
                 strictMask,
                 by = c('Chr','Pos')) %>%
  mutate(strict = "pass")

Reg_mask <- rbind(Reg_strict, Reg_not_strict)

#Get P-Values and significant sites
Reg_mask$plog10 <- -log10(pchisq(Reg_mask$deviance, 1, 
                            lower.tail = FALSE))
Reg_mask<-Reg_mask[which(Reg_mask$plog10 >= 6),]
#Count number of populations where a variant is significant
NumPop <- as.data.frame(table(Reg_mask$Chr,Reg_mask$Pos))
names(NumPop) <- c('Chr','Pos', 'SigHits')
NumPop$Chr<-as.numeric(as.character(NumPop$Chr)) 
NumPop$Pos<-as.numeric(as.character(NumPop$Pos))

SinglePop <- NumPop[which(NumPop$SigHits==1),]
SinglePop <- merge(Reg_mask,
                   SinglePop, 
                   by = c('Chr','Pos')) %>%
  group_by(Population, strict) %>%
  summarize(single_count = n())

DoublePop <- NumPop[which(NumPop$SigHits>1),]
DoublePop <- merge(Reg_mask,
                   DoublePop, 
                   by = c('Chr','Pos')) %>%
  group_by(Population, strict) %>%
  summarize(double_count = n())

Combined_count <- merge(DoublePop, SinglePop, by = c('Population', 'strict'))

Combined_count[which(Combined_count$Pop%in% phase1),]$Population <- paste(Combined_count[which(Combined_count$Pop%in% phase1),]$Population, '*')

#save to file
saveRDS(Combined_count,
            paste0(DataPath,
                   'overLap.rds'))

pos_count <- Reg_mask %>%
  group_by(Chr,Pos,strict) %>%
  summarize(count = n()) %>%
  group_by(count, strict) %>%
  summarize(freq = n())

#save to file
saveRDS(pos_count,
            paste0(DataPath,
                   'pos_count.rds'))
```
